Image projections along an axe

### 2006-01-22 article/Article.tex
write the article
darcs-hash:20060122220909-2fc9d-5101590b3bd35727465dc5294e67e16f6420a655.gz

% The title should be descriptive enough for people to be able to find
% the relevant document. 
\title{}

% Increment the release number whenever significant changes are made.
% The author and/or editor can define 'significant' however they like.
\release{0.00}

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{}
\authoraddress{}

@@

\begin{abstract}
\noindent
% The abstract should be a paragraph or two long, and describe the
% scope of the document.
\end{abstract}

\tableofcontents

% \url{http://www.itk.org}
% \code{Insight/Documentation/Style.pdf}
% \section{Principles of Solar Spot Detection}
% \cite{ITKSoftwareGuide}.
% \doxygen{ImageToImageFilter}

% \small \begin{verbatim}
% \end{verbatim} \normalsize

% The {itemize} environment uses a bullet for each \item.  If you want the 
% \item's numbered, use the {enumerate} environment instead.
% \begin{itemize}
%   \item  Insight Toolkit 2.4.
%   \item  CMake 2.2
% \end{itemize}

% \ref{cthead1}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{maximum0}
\caption{Maximum projection, axe=0.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{maximum1}
\caption{Maximum projection, axe=1.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{maximum}
\caption{Maximum projection, axe=2.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{minimum}
\caption{Minimum projection}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{mean}
\caption{Mean projection.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{median}
\caption{Median projection}
\end{figure}

% \begin{figure}[htbp]
% \centering
% \includegraphics[scale=0.75]{sum}
% \caption{Sum projection}
% \end{figure}

>>>>

% The title should be descriptive enough for people to be able to find
% the relevant document. 
\title{Image projections along an axe}

% Increment the release number whenever significant changes are made.
% The author and/or editor can define 'significant' however they like.
% \release{0.00}

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Ga\"etan Lehmann}
\authoraddress{Unit\'e de Biologie du D\'eveloppement et de la Reproduction, Institut National de la Recherche Agronomique, 78350 Jouy-en-Josas, France}

@@

\begin{abstract}
\noindent
Image projection is a very common task in image analysis to reduce the dimension
of an image. A base filter is provided with some specialized filters which
implement different projection methods.
\end{abstract}

% \tableofcontents

\section{Introduction}
ITK already provide a class to project an image along an axe:
\doxygen{AccumulateImageFilter}. However, this class is limited to sum
projection and mean projection.

\section{Implementation}

The base class, ProjectionImageFilter, is mainly a copy of
AccumulateImageFilter, with some enhancement:
\begin{itemize}
  \item projection is delegated to a functor
  \item report its progress
\end{itemize}
Different projection types are provided, with specialized classes:
\begin{itemize}
  \item maximum projection (MaximumProjectionImageFilter)
  \item minimum projection (MinimumProjectionImageFilter)
  \item mean projection (MeanProjectionImageFilter)
  \item median projection (MedianProjectionImageFilter)
  \item sum projection (SumProjectionImageFilter)
\end{itemize}
Other projections types can easily be created just by providing a new functor.

AccumulateImageFilter is not modifiable to implement functor support without
breaking the API.

\section{Performance}

A timing test comparing performance on a $371 \times 371 \times 34$
image showed that the results are similar with AccumulateImageFilter
and the equivalent ProjectionImageFilter, and that the new projection
types have similar performance. The results achieved on an Athlon 64 
Processor 2800+ (1802 MHz) with 512Kb cache, 512 Mb of RAM and gcc
4.0.2 are shown in Table~\ref{perf}.

\begin{table}[htbp]
\centering
\begin{tabular}{cc}
\hline
Filter & execution time \\
\hline
\hline
MaximumProjectionImageFilter & 0.652 s \\
MinimumProjectionImageFilter  & 0.614 s\\
SumProjectionImageFilter  & 0.614 s\\
MeanProjectionImageFilter  & 0.619 s\\
MedianProjectionImageFilter  & 0.831 s\\
AccumulateImageFilter, sum  & 0.615 s\\
AccumulateImageFilter, mean  & 0.617 s\\
\hline
\end{tabular}
\caption{Execution time.\label{perf}}
\end{table}

\section{Examples}
Figure~\ref{maximum0}, Figure~\ref{maximum1}, Figure~\ref{maximum}, Figure~\ref{minimum}, Figure~\ref{mean} and Figure~\ref{median} show the result of different projection on different axes. The result of SumProjectionImageFilter in similar to the result of MeanProjectionImageFilter, but with a higher range of values, and so is not shown here.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{maximum0}
\caption{MaximumProjectionImageFilter, projection along axe x.\label{maximum0}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{maximum1}
\caption{MaximumProjectionImageFilter, projection along axe y.\label{maximum1}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{maximum}
\caption{MaximumProjectionImageFilter, projection along axe z.\label{maximum}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{minimum}
\caption{MinimumProjectionImageFilter, projection along axe z.\label{minimum}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{mean}
\caption{MeanProjectionImageFilter, projection along axe z.\label{mean}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.75]{median}
\caption{MedianProjectionImageFilter, projection along axe z\label{median}}
\end{figure}

\section{Sample code}

\small \begin{verbatim}
#include "itkImageFileReader.h"
#include "itkImageFileWriter.h"
#include "itkCommand.h"
#include "itkSimpleFilterWatcher.h"

#include "itkMaximumProjectionImageFilter.h"
#include "itkExtractImageFilter.h"

int main(int, char * argv[])
{
  int axe = atoi(argv[1]);

  typedef unsigned char PType;
  typedef itk::Image< PType, 3 > IType;
  typedef itk::Image< PType, 2 > IType2;

  typedef itk::ImageFileReader< IType > ReaderType;
  ReaderType::Pointer reader = ReaderType::New();
  reader->SetFileName( argv[2] );

  typedef itk::MaximumProjectionImageFilter< IType, IType > FilterType;
  FilterType::Pointer filter = FilterType::New();
  filter->SetInput( reader->GetOutput() );
  filter->SetAxe( axe );

  itk::SimpleFilterWatcher watcher(filter, "filter");

  filter->Update();

  IType::SizeType inputSize = filter->GetOutput()->GetLargestPossibleRegion().GetSize();

  typedef itk::ExtractImageFilter< IType, IType2 > ExtractType;
  ExtractType::Pointer extract = ExtractType::New();
  extract->SetInput( filter->GetOutput() );

  IType::SizeType size;
  for(int i=0; i<=3; i++) {
    if(i == axe) {
      size[i] = 0;
    } else {
      size[i] = inputSize[i];
     }
  }
  IType::IndexType idx;
  idx.Fill(0);

  IType::RegionType region;
  region.SetSize( size );
  region.SetIndex( idx );
  extract->SetExtractionRegion( region );

  typedef itk::ImageFileWriter< IType2 > WriterType;
  WriterType::Pointer writer = WriterType::New();
  writer->SetInput( extract->GetOutput() );
  writer->SetFileName( argv[3] );
  writer->Update();

  return 0;
}
\end{verbatim} \normalsize

\section{Conclusion}
The new classes give more feature and more flexibility with equivalent
performance than the current available filter, but loss the compatibility with
the previous filter. I propose to mark AccumulateImageFilter as deprecated in
the next release of ITK to incite user to the new class.

\section{Acknowledgments}
We thank Dr Pierre Adenot and MIMA2 confocal facilities
(\url{http://mima2.jouy.inra.fr})
for providing image samples.
